Cannibalism among the supermassive black holes

### 2018-08-10 paper.tex
Added more edits from Jerry.  Added Merger Summary.

\documentclass[english, apj]{emulateapj}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{rotating}
\usepackage{units}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{url}
\usepackage{babel}
%\usepackage{siunitx}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage[all]{hypcap}

@@

\section{Introduction}\label{sec:introduction}
What happens to the massive black holes centrally located in galaxies when those galaxies merge with others?  There are three possibilities: they merge, they are ejected, or they remain in orbit in the merged galaxy.  Combinations of these options are also possible with the spin induced kick at merger leading to an ejection or extended orbit of the merged black hole.  We now know from pulsar timing measurements (ADD REFERENCES) that the gravitational wave background is probably too low for most galaxy mergers (mass-weighted) (BUT SEE MIDLETON ET AL 2018) to lead to BH mergers, but the outcome remains uncertain.  We will argue in this paper that a very common outcome--especially for the lower mass black holes--is for the injected objects to remain as orbiting X-ray sources in normal massive galaxies waiting to be identified by current observational techniques.

A majority of all galaxies with masses $>10^8\msun$ at $z=0$ are believed to host super-massive black holes (SMBHs) in their cores.  What is the ultimate fate of the supermassive black holes (SMBH) resident at the centers of these massive galaxies?  For the galaxies themselves we know that mergers are frequent from the observed (VAN DOKKUM PAPERS??) evolution of the size and mass of these systems.  But assuming that all galaxy mergers lead to SMBH mergers overpredicts the observations of gravitational waves from the pulsar timing arrays (PTAs) (\citet{2015ApJ...799..178K}, \citet{2014ApJ...789..156M}, \citet{2008MNRAS.390..192S}, \citet{2013MNRAS.433L...1S}, \citet{2018ApJ...856...42S}, \citet{2009MNRAS.394.2255S}, \citet{2018arXiv180403143I}, NATARAJAN??).  But we do not often see multiple SMBHs at the centers of massive ellipticals.  So, what does happen?  One paper (\citet{2018MNRAS.473.3410R}) has indicated that dynamical interactions among multiple black holes, which eject a non-negligible fraction of the mass, may solve this problem.  The present paper also addresses this purported dynamical solution, focusing attention on the large fraction of lower mass black holes that remain to be detected as they orbit in massive galaxies.  What we argue in this paper is that as a result of the N-body interactions among the infalling black holes some are ejected, some remain in extended orbits and some (the few most massive ones) do in fact merge but we do not exceed the PTA limits, and the orbiting ones should be detectable via their ability to accrete gas and emit radiation.

Here we look at the merger history of 12 exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs diffuse into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with $10^{12} \msun < M < 10^{13} \msun$. Galaxies with lower masses have too few mergers with SMBH hosting galaxies. Galaxies with higher masses are more extended, making dynamical friction processes less efficient and hence failing to drive SMBHs into the host galaxy core.

@@

\section{Results}\label{sec:results}
\begin{table*}
\centering
\caption{SMBH statistics}
\begin{tabular}{c| c |c |c}
Galaxy & Infalling SMBHs & SMBHs with $t_{fric}<100\,t_H$ & Mergers \\
\hline
A & 206 & 19 & 4 \\
B & 29 & 12 & 4 \\
C & 3 & 3 & 2 \\
D & 4 & 4 & 2 \\
\end{tabular}
\end{table*}

>>>>

\documentclass[english, apj]{emulateapj}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{rotating}
\usepackage{units}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{url}
\usepackage{babel}
\usepackage{float}
%\usepackage{siunitx}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage[all]{hypcap}

@@

\section{Introduction}\label{sec:introduction}
What happens to the massive black holes centrally located in galaxies when those galaxies merge with others?  There are three possibilities: they merge, they are ejected, or they remain in orbit in the merged galaxy.  Combinations of these options are also possible with the spin induced kick at merger leading to an ejection or extended orbit of the merged black hole.  We now know from pulsar timing measurements (ADD REFERENCES) that the gravitational wave background is probably too low for most galaxy mergers (mass-weighted) (BUT SEE MIDDLETON ET AL 2018) to lead to BH mergers, but the outcome remains uncertain.  We will argue in this paper that a very common outcome--especially for the lower mass black holes--is for the injected objects to remain as orbiting X-ray sources in normal massive galaxies waiting to be identified by current observational techniques.

A majority of all galaxies with masses $>10^8\msun$ at $z=0$ are believed to host super-massive black holes (SMBHs) in their cores.  What is the ultimate fate of the supermassive black holes (SMBH) resident at the centers of these massive galaxies?  For the galaxies themselves we know that mergers are frequent from the observed (VAN DOKKUM PAPERS??) evolution of the size and mass of these systems.  But assuming that all galaxy mergers lead to SMBH mergers overpredicts the observations of gravitational waves from the pulsar timing arrays (PTAs) (\citet{2015ApJ...799..178K}, \citet{2014ApJ...789..156M}, \citet{2008MNRAS.390..192S}, \citet{2013MNRAS.433L...1S}, \citet{2018ApJ...856...42S}, \citet{2009MNRAS.394.2255S}, \citet{2018arXiv180403143I}).  But we do not often see multiple SMBHs at the centers of massive ellipticals.  So, what does happen?  One paper (\citet{2018MNRAS.473.3410R}) has indicated that dynamical interactions among multiple black holes, which eject a non-negligible fraction of the mass, may solve this problem.  The present paper also addresses this purported dynamical solution, focusing attention on the large fraction of lower mass black holes that remain to be detected as they orbit in massive galaxies.  What we argue in this paper is that as a result of the N-body interactions among the infalling black holes some are ejected, some remain in extended orbits and some (the few most massive ones) do in fact merge but we do not exceed the PTA limits, and the orbiting ones should be detectable via their ability to accrete gas and emit radiation.

Here we look at the merger history of 13 exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs diffuse into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with $10^{12} \msun < M < 10^{13} \msun$. Galaxies with lower masses have too few mergers with SMBH hosting galaxies. Galaxies with higher masses are more extended, making dynamical friction processes less efficient and hence failing to drive SMBHs into the host galaxy core.

@@

\section{Results}\label{sec:results}
\begin{figure}[H]
\begin{center}
\includegraphics[width=0.45\textwidth]{plots/Summary_Results_All_Galaxies.png}
\caption{Histogram of merged, ejected, and orbiting SMBHs}
\label{fig:meosmbh}
\end{center}
\end{figure}

\begin{table*}
\centering
\caption{SMBH statistics}
\begin{tabular}{c| c |c |c}
Galaxy & Infalling SMBHs & SMBHs with $t_{fric}<100\,t_H$ & Mergers \\
\hline
A & 206 & 19 & 4 \\
B & 29 & 12 & 4 \\
C & 3 & 3 & 2 \\
D & 4 & 4 & 2 \\
\end{tabular}
\end{table*}
